---
title: "Columbus Public Health"
author: "Aung Myo Htut"
date: "2024-02-06"
output:
  
  html_document: default
  always_allow_html: true
  latex_engine: xelatex
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the library
```{r load library, warning = FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tools)
library(openxlsx)
library(maps)
library(sf)
library(labelled)
library(table1)
library(kableExtra)
library(knitr)
library(webshot)
```

#SVI and Select variables of interest
```{r SVI}
svi <- read_csv("CDC_2020_SVI_Connecticut_County_Level.csv", show_col_types = FALSE)

#Select Variables of Interest out of 158 Variables
cdc_svi <- svi %>% 
  select("COUNTY", "E_TOTPOP", "E_HU", "E_HH", "E_POV150", "E_UNEMP","E_NOHSDP",
         "E_UNINSUR", "E_AGE65", "E_AGE17", "E_DISABL", "E_SNGPNT", "E_MINRTY")

#Rename variables
names(cdc_svi) <- c("county", "total_pop", "housing_units", "households", "belowpov150", "unemployment",
                    "nohsdp", "uninsured", "age65","age17", "disablity", "single_parent", "race_minority")

```

#Handling Shapefile Data for Connecticut Counties
```{r shapefile, warning=FALSE}
#Save original working directory
original_wd <- getwd()

#Change working directory for this task
setwd("C:/Users/draun/Desktop/Columbus Public Health/cb_2018_us_county_500k")

#Load shape file
my_shape <- st_read("C:/Users/draun/Desktop/Columbus Public Health/cb_2018_us_county_500k/cb_2018_us_county_500k.shp")

# Define the list of county names
counties <- c("Fairfield", "Hartford", "Litchfield", "Middlesex", "New Haven", "New London", "Tolland", "Windham")

# Filter the dataset
filtered_dataset <- my_shape %>%
                    filter(NAME %in% counties) %>%
                    filter (STATEFP %in% "09")



#Rename variables
shapefile <- filtered_dataset %>%
  rename(
    "statefp" = "STATEFP",
    "countyfp" = "COUNTYFP",
    "countyns" = "COUNTYNS",
    "affgeoid" = "AFFGEOID",
    "geoid" = "GEOID",
    "county" = "NAME",
    "lsad" = "LSAD",
    "aland" = "ALAND",
    "awater" = "AWATER",
    "geometry" = "geometry"

  )

#Revert back to original working directory
setwd(original_wd)

# #Export to shapefile
# st_write(shapefile,"C:/Users/draun/Desktop/Columbus Public Health/Shapefile/my_shapefile.shp")
```

#Load Death Data and Aggregate at County Level
```{r death data}
#Load the data
death_data <- read_csv("Connecticut_Accidental_Drug_Related_Deaths_2012-2022.csv", show_col_types = F)

#Filter Data
death_data <-death_data[,3:9]

#Rename Columns
names(death_data) <- c("age", "sex", "race","ethnicity", "city", "county", "state")

#Filter only people from Connecticut
death_filter <- death_data %>% filter(state == "CT")

# Change all capital cases To Lower case observations
death_filter$county <- toTitleCase(tolower(death_filter$county))
head(death_filter)

#Aggregate the death data by county
death_data_aggregate <- death_filter %>%
  group_by(county) %>%
  summarise(total_deaths = n())


# Remove observations with specified county names and NA values
new_death <- death_data_aggregate[!(death_data_aggregate$county %in% c("Middlesez", "Newlondon", "Plymouth") | is.na(death_data_aggregate$county)), ]

# View the modified dataset
head(new_death, n = 8)

```
#Join the Datasets
```{r data joining}
#Joining Data
combined_data <- left_join(cdc_svi, new_death, by = "county")
head(combined_data)
```

#Exporting to Excel File
```{r Exporting Data}
#Exporting Data
#write.xlsx(combined_data,"data.xlsx")
```

#Merge the datab and create table
```{r merging, warning=FALSE}
# Merge the shapefile data with the merged_data
map_data <- merge(shapefile, combined_data, by = "county")

#labeling variables
var_label(map_data) <-
  list(county = "County",
       statefp = "State FP Codes",
       countyfp = "County FP Codes",
       countyns = "Countyns",
       affgeoid = "Affgeoid",
       geoid = "Geo ID",
       lsad = "lsad",
       aland = "aland",
       awater = "awater",
       total_pop = "Total Population",
       housing_units = "Housing Units Estimate",
       households = "Households Estimate",
       belowpov150 = "Persons Below 150% Poverty Estimate",
       unemployment = "Civilian (age+) unemployed estimate",
       nohsdp = "Persons (age25+) with no high school diploma",
       uninsured = "Uninsured population",
       age65 = "Persons aged 65 and older",
       age17 = "Persons aged 17 and younger",
       disablity = "Population with disability",
       single_parent = "Single parent household with children < 18",
       race_minority = "Minority Race",
       total_deaths = "Total Deaths",
       geometry = "Geometry")

#Table
table <- table1(~ total_pop + housing_units + households + belowpov150 + unemployment + nohsdp +
         uninsured + age65 + age17 + disablity + single_parent + race_minority + 
         total_deaths, data = map_data, overall = "Total")
#table

#Format Table
kable_table <- kable(table, format = "html", row.names = FALSE,escape = TRUE)

format_table <- kable_table %>% 
  kable_styling(full_width = FALSE) %>%
  row_spec(row = 0, color = "white", background = "#FF5733", font_size = 15)%>% #for title
  row_spec(row = 2, bold = TRUE, font_size = 13) %>% #bold each variable
  row_spec(row = 5, bold = TRUE, font_size = 13) %>%
  row_spec(row = 8, bold = TRUE, font_size = 13) %>%
  row_spec(row = 11, bold = TRUE, font_size = 13) %>%
  row_spec(row = 14, bold = TRUE, font_size = 13) %>%
  row_spec(row = 17, bold = TRUE, font_size = 13) %>%
  row_spec(row = 20, bold = TRUE, font_size = 13) %>%
  row_spec(row = 23, bold = TRUE, font_size = 13) %>%
  row_spec(row = 26, bold = TRUE, font_size = 13) %>%
  row_spec(row = 29, bold = TRUE, font_size = 13) %>%
  row_spec(row = 32, bold = TRUE, font_size = 13) %>%
  row_spec(row = 35, bold = TRUE, font_size = 13) %>%
  row_spec(row = 38, bold = TRUE, font_size = 13)

format_table
```

#Choropleth County and Unemployment Estimate
```{r map, warning=FALSE}
#Plotting
ggplot(map_data) +
  geom_sf(aes(fill = unemployment)) +
  geom_sf_text(aes(label = county), size = 3, color = "darkred") +
  labs(title = "Unemployment Estimates in the State of Conneticut",
       fill = "Unemployment") +
  scale_fill_gradient(low = "lightcyan", high = "deepskyblue4") +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))
```

#Choropleth Map of Deaths in Counties
```{r, warning=FALSE}
# Create a new column for labels
map_data$label <- paste(map_data$county, map_data$total_deaths, sep = ": ")

# Plotting
ggplot(map_data) +
  geom_sf(aes(fill = total_deaths)) +
  geom_sf_text(aes(label = label), size = 3, color = "orange") +  # Corrected this line
  labs(title = "Drug Related Total Deaths in Connecitcut", fill = "Deaths") +
  scale_fill_gradient(low = "lightcyan", high = "dodgerblue4") +
  theme_void() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))

```

#Poverty Vs Total Death Scatter Plot
```{r Poverty vs death}
#I know this wouldn't look good because of too few points
#There is still positive relation though
ggplot(map_data, aes(x = belowpov150, y = total_deaths)) +
  geom_point() +
  labs(x = "Population Below 150% of Poverty Line",
       y = "Drug Related Total Deaths",
       title = "Poverty vs. Total Deaths by County")

```

#Bar Chart of Minority Population:
```{r}
ggplot(map_data, aes(x = county, y = race_minority)) +
  geom_col(fill = "darkolivegreen4")+
  geom_text(aes(label = race_minority), vjust = -0.5, color = "red", size = 2.5) +
  scale_y_continuous(breaks = c(seq(0,400000,40000))) +
  labs(x = "County",
       y = "Minority Population",
       title = "Minority Population by County in Conneticut") +
  theme_classic() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))

```

